---
title: 'Voting rights and the HR2W: Initial analysis and next steps brainstorm'
author: "Kristin Dobbin"
date: "2023-07-05"
output:
  html_document: default
  pdf_document: default
knit: (function(rmdfile, ...) { rmarkdown::render(rmdfile, output_dir=here::here("Docs"))})
---

```{r setup, include=FALSE, echo=FALSE}
#markdown settings
knitr::opts_knit$set(root.dir = "/Users/KristinDobbin/Library/CloudStorage/Dropbox/Projects/R_Projects/Votingrights_-_HR2W", warning=FALSE, message=FALSE, include=FALSE)

#load libraries
library(tidyverse)
```

```{r data, include=FALSE, echo=FALSE}
###Load data and initial manipulations

CWSgov <- read.csv(here::here("Data_raw/FINAL 2023 Inst. type update.csv"))
CWSgov <- CWSgov %>% filter(Primacy_FINAL != "EPA Region 9") #for purposes of this project get rid of EPA regulated Tribal water systems
CWSgov$Primacy_FINAL <- as.factor(CWSgov$Primacy_FINAL)
CWSgov$Primacy_FINAL <- droplevels(CWSgov$Primacy_FINAL)
CWSgov <- CWSgov[,c(1:4,10:11)]
CWSgov$PWSID <- as.factor(CWSgov$PWSID)
CWSgov$Counties_Served <- as.factor(CWSgov$Counties_Served)
CWSgov$Final_inst_update <- as.factor(CWSgov$Final_inst_update)
CWSgov$Inst_Subtype  <- as.factor(CWSgov$Inst_Subtype)

### create new variables
## create ancillary variable
CWSgov$ancillary <- ifelse(CWSgov$Final_inst_update == "Community Services District" | CWSgov$Final_inst_update == "County Service Area" | CWSgov$Final_inst_update == "Municipal Water District" | CWSgov$Final_inst_update == "City" | CWSgov$Final_inst_update == "County Sanitation District"| CWSgov$Final_inst_update == "County Water District"| CWSgov$Final_inst_update == "County Waterworks District" | CWSgov$Final_inst_update == "Irrigation District"| CWSgov$Final_inst_update == "Maintenance District"| CWSgov$Final_inst_update == "Municipal Utility District"| CWSgov$Final_inst_update == "Public Utility District"| CWSgov$Final_inst_update == "Resort Improvement District"| CWSgov$Final_inst_update == "Resource Conservation District"| CWSgov$Final_inst_update == "Sanitary District"| CWSgov$Final_inst_update == "Water Conservation District"| CWSgov$Final_inst_update == "Special Act District" | CWSgov$Final_inst_update == "Mutual Benefit" | CWSgov$Final_inst_update == "Joint Powers Authority/Agreement" | CWSgov$Final_inst_update == "Investor Owned Utility" | CWSgov$Final_inst_update == "California Water District","No", NA)

CWSgov$ancillary <- ifelse(CWSgov$Final_inst_update == "County" | CWSgov$Final_inst_update == "Federal" | CWSgov$Final_inst_update == "Private - Ancillary" | CWSgov$Final_inst_update == "School District" | CWSgov$Final_inst_update == "State" | CWSgov$Final_inst_update == "Mobile Home Park" , "Yes", CWSgov$ancillary)

## create public/private variable
CWSgov$public <- ifelse(CWSgov$Final_inst_update == "Community Services District" | CWSgov$Final_inst_update == "County Service Area" | CWSgov$Final_inst_update == "Municipal Water District" | CWSgov$Final_inst_update == "City" | CWSgov$Final_inst_update == "County Sanitation District"| CWSgov$Final_inst_update == "County Water District"| CWSgov$Final_inst_update == "County Waterworks District" | CWSgov$Final_inst_update == "Irrigation District"| CWSgov$Final_inst_update == "Maintenance District"| CWSgov$Final_inst_update == "Municipal Utility District"| CWSgov$Final_inst_update == "Public Utility District"| CWSgov$Final_inst_update == "Resort Improvement District"| CWSgov$Final_inst_update == "Resource Conservation District"| CWSgov$Final_inst_update == "Sanitary District"| CWSgov$Final_inst_update == "Water Conservation District"| CWSgov$Final_inst_update == "Special Act District" | CWSgov$Final_inst_update == "Mutual Benefit" | CWSgov$Final_inst_update == "Joint Powers Authority/Agreement" | CWSgov$Final_inst_update == "California Water District" | CWSgov$Final_inst_update == "County" | CWSgov$Final_inst_update == "Federal" | CWSgov$Final_inst_update == "School District" | CWSgov$Final_inst_update == "State" | CWSgov$Final_inst_update == "Tribal", "Yes", "No")

## create voting enfranchisement variable
CWSgov$voting <- ifelse(CWSgov$Final_inst_update == "Community Services District" | CWSgov$Final_inst_update == "County Service Area" | CWSgov$Final_inst_update == "Municipal Water District" | CWSgov$Final_inst_update == "City" | CWSgov$Final_inst_update == "County Sanitation District"| CWSgov$Final_inst_update == "County Water District"| CWSgov$Final_inst_update == "County Waterworks District" | CWSgov$Final_inst_update == "Irrigation District"| CWSgov$Final_inst_update == "Maintenance District"| CWSgov$Final_inst_update == "Municipal Utility District"| CWSgov$Final_inst_update == "Public Utility District"| CWSgov$Final_inst_update == "Resort Improvement District"| CWSgov$Final_inst_update == "Resource Conservation District"| CWSgov$Final_inst_update == "Sanitary District"| CWSgov$Final_inst_update == "Water Conservation District", "Full", NA)

CWSgov$voting <- ifelse(CWSgov$Final_inst_update == "Mutual Benefit" | CWSgov$Final_inst_update == "Irrigation District","Limited", CWSgov$voting)

CWSgov$voting <- ifelse(CWSgov$Final_inst_update == "Investor Owned Utility", "None", CWSgov$voting)

CWSgov$voting <- ifelse(CWSgov$Final_inst_update == "California Water District" | CWSgov$Final_inst_update == "Joint Powers Authority/Agreement" | CWSgov$Final_inst_update == "Special Act District" , "Variable", CWSgov$voting)

#create second version of voting variable including MHPs as "none"
CWSgov$voting2 <- ifelse(CWSgov$Final_inst_update == "Community Services District" | CWSgov$Final_inst_update == "County Service Area" | CWSgov$Final_inst_update == "Municipal Water District" | CWSgov$Final_inst_update == "City" | CWSgov$Final_inst_update == "County Sanitation District"| CWSgov$Final_inst_update == "County Water District"| CWSgov$Final_inst_update == "County Waterworks District" | CWSgov$Final_inst_update == "Irrigation District"| CWSgov$Final_inst_update == "Maintenance District"| CWSgov$Final_inst_update == "Municipal Utility District"| CWSgov$Final_inst_update == "Public Utility District"| CWSgov$Final_inst_update == "Resort Improvement District"| CWSgov$Final_inst_update == "Resource Conservation District"| CWSgov$Final_inst_update == "Sanitary District"| CWSgov$Final_inst_update == "Water Conservation District", "Full", NA)

CWSgov$voting2 <- ifelse(CWSgov$Final_inst_update == "Mutual Benefit" | CWSgov$Final_inst_update == "Irrigation District","Limited", CWSgov$voting2)

CWSgov$voting2 <- ifelse(CWSgov$Final_inst_update == "Investor Owned Utility" | CWSgov$Final_inst_update == "Mobile Home Park", "None", CWSgov$voting2)

CWSgov$voting2 <- ifelse(CWSgov$Final_inst_update == "California Water District" | CWSgov$Final_inst_update == "Joint Powers Authority/Agreement" | CWSgov$Final_inst_update == "Special Act District" , "Variable", CWSgov$voting2)

#create third version of voting variable including all ancillary systems as "none"
CWSgov$voting3 <- ifelse(CWSgov$Final_inst_update == "Community Services District" | CWSgov$Final_inst_update == "County Service Area" | CWSgov$Final_inst_update == "Municipal Water District" | CWSgov$Final_inst_update == "City" | CWSgov$Final_inst_update == "County Sanitation District"| CWSgov$Final_inst_update == "County Water District"| CWSgov$Final_inst_update == "County Waterworks District" | CWSgov$Final_inst_update == "Irrigation District"| CWSgov$Final_inst_update == "Maintenance District"| CWSgov$Final_inst_update == "Municipal Utility District"| CWSgov$Final_inst_update == "Public Utility District"| CWSgov$Final_inst_update == "Resort Improvement District"| CWSgov$Final_inst_update == "Resource Conservation District"| CWSgov$Final_inst_update == "Sanitary District"| CWSgov$Final_inst_update == "Water Conservation District", "Full", NA)

CWSgov$voting3 <- ifelse(CWSgov$Final_inst_update == "Mutual Benefit" | CWSgov$Final_inst_update == "Irrigation District","Limited", CWSgov$voting3)

CWSgov$voting3 <- ifelse(CWSgov$Final_inst_update == "Investor Owned Utility" | CWSgov$Final_inst_update == "Mobile Home Park" | CWSgov$Final_inst_update == "County" | CWSgov$Final_inst_update == "Federal" | CWSgov$Final_inst_update == "Private - Ancillary" | CWSgov$Final_inst_update == "School District" | CWSgov$Final_inst_update == "State", "None", CWSgov$voting3)

CWSgov$voting3 <- ifelse(CWSgov$Final_inst_update == "California Water District" | CWSgov$Final_inst_update == "Joint Powers Authority/Agreement" | CWSgov$Final_inst_update == "Special Act District" , "Variable", CWSgov$voting3)
##mutate to create new variables following initial script but instead of NA for variable, use variable and can filter out later but that way can check with Camille and Justin and discuss what to do

## create a fourth voting enfranchisement variable that is more like the first but calls out MHPs as none isntead of NA
CWSgov$voting4 <- ifelse(CWSgov$Final_inst_update == "Community Services District" | CWSgov$Final_inst_update == "County Service Area" | CWSgov$Final_inst_update == "Municipal Water District" | CWSgov$Final_inst_update == "City" | CWSgov$Final_inst_update == "County Sanitation District"| CWSgov$Final_inst_update == "County Water District"| CWSgov$Final_inst_update == "County Waterworks District" | CWSgov$Final_inst_update == "Irrigation District"| CWSgov$Final_inst_update == "Maintenance District"| CWSgov$Final_inst_update == "Municipal Utility District"| CWSgov$Final_inst_update == "Public Utility District"| CWSgov$Final_inst_update == "Resort Improvement District"| CWSgov$Final_inst_update == "Resource Conservation District"| CWSgov$Final_inst_update == "Sanitary District"| CWSgov$Final_inst_update == "Water Conservation District", "Full", NA)

CWSgov$voting4 <- ifelse(CWSgov$Final_inst_update == "Mutual Benefit" | CWSgov$Final_inst_update == "Irrigation District","Limited", CWSgov$voting4)

CWSgov$voting4 <- ifelse(CWSgov$Final_inst_update == "Investor Owned Utility" | CWSgov$Final_inst_update == "Mobile Home Park", "None", CWSgov$voting4)

CWSgov$voting4 <- ifelse(CWSgov$Final_inst_update == "California Water District" | CWSgov$Final_inst_update == "Joint Powers Authority/Agreement" | CWSgov$Final_inst_update == "Special Act District" , "Variable", CWSgov$voting4)


## create final variable about whether there is a board and if so who is elgible to serve on it
#create third version of voting variable including all ancillary systems as "none"
CWSgov$board <- ifelse(CWSgov$Final_inst_update == "Community Services District" | CWSgov$Final_inst_update == "County Service Area" | CWSgov$Final_inst_update == "Municipal Water District" | CWSgov$Final_inst_update == "City" | CWSgov$Final_inst_update == "County Sanitation District"| CWSgov$Final_inst_update == "County Water District"| CWSgov$Final_inst_update == "County Waterworks District" | CWSgov$Final_inst_update == "Irrigation District"| CWSgov$Final_inst_update == "Maintenance District"| CWSgov$Final_inst_update == "Municipal Utility District"| CWSgov$Final_inst_update == "Public Utility District"| CWSgov$Final_inst_update == "Resort Improvement District"| CWSgov$Final_inst_update == "Resource Conservation District"| CWSgov$Final_inst_update == "Sanitary District"| CWSgov$Final_inst_update == "Water Conservation District", "Full", NA)

CWSgov$board <- ifelse(CWSgov$Final_inst_update == "California Water District" | CWSgov$Final_inst_update == "Mutual Benefit" | CWSgov$Final_inst_update == "Irrigation District","Limited", CWSgov$board)

CWSgov$board <- ifelse(CWSgov$Final_inst_update == "Investor Owned Utility" | CWSgov$Final_inst_update == "Mobile Home Park" | CWSgov$Final_inst_update == "County" | CWSgov$Final_inst_update == "Federal" | CWSgov$Final_inst_update == "Private - Ancillary" | CWSgov$Final_inst_update == "School District" | CWSgov$Final_inst_update == "State", "None", CWSgov$board)

CWSgov$board <- ifelse( CWSgov$Final_inst_update == "Joint Powers Authority/Agreement" | CWSgov$Final_inst_update == "Special Act District" | CWSgov$Final_inst_update == "County Sanitation District" | CWSgov$Final_inst_update == "Irrigation District", "Variable", CWSgov$board)

CWSgov$boardselection <- ifelse(CWSgov$Final_inst_update == "Community Services District" |  CWSgov$Final_inst_update == "Municipal Water District" | CWSgov$Final_inst_update == "City" | CWSgov$Final_inst_update == "County Sanitation District"| CWSgov$Final_inst_update == "County Water District"| CWSgov$Final_inst_update == "Maintenance District"| CWSgov$Final_inst_update == "Municipal Utility District"| CWSgov$Final_inst_update == "Public Utility District"| CWSgov$Final_inst_update == "Resort Improvement District"| CWSgov$Final_inst_update == "Resource Conservation District"| CWSgov$Final_inst_update == "Sanitary District"| CWSgov$Final_inst_update == "Water Conservation District", "Board selected by residents", NA)

CWSgov$boardselection <- ifelse(CWSgov$Final_inst_update == "Mutual Benefit","Board selected by limited residents", CWSgov$boardselection)

CWSgov$boardselection <- ifelse(CWSgov$Final_inst_update == "Investor Owned Utility" | CWSgov$Final_inst_update == "County Service Area" | CWSgov$Final_inst_update == "County Waterworks District" | CWSgov$Final_inst_update == "Mobile Home Park" | CWSgov$Final_inst_update == "County" | CWSgov$Final_inst_update == "Federal" | CWSgov$Final_inst_update == "Private - Ancillary" | CWSgov$Final_inst_update == "School District" | CWSgov$Final_inst_update == "State", "Board not selected by residents", CWSgov$boardselection)

CWSgov$boardselection <- ifelse(CWSgov$Final_inst_update == "Mobile Home Park" | CWSgov$Final_inst_update == "Federal" | CWSgov$Final_inst_update == "Private - Ancillary" | CWSgov$Final_inst_update == "School District" | CWSgov$Final_inst_update == "State", "No board", CWSgov$boardselection)

CWSgov$boardselection <- ifelse(CWSgov$Final_inst_update == "California Water District" | CWSgov$Final_inst_update == "Irrigation District"| CWSgov$Final_inst_update == "Joint Powers Authority/Agreement" | CWSgov$Final_inst_update == "Special Act District" , "Variable", CWSgov$boardselection)

#change to factors
CWSgov$ancillary <- as.factor(CWSgov$ancillary)
CWSgov$public <- as.factor(CWSgov$public)
CWSgov$voting <- as.factor(CWSgov$voting)
CWSgov$voting2 <- as.factor(CWSgov$voting2)
CWSgov$voting3 <- as.factor(CWSgov$voting3)
CWSgov$boardselection <- as.factor(CWSgov$boardselection)

## add in outcome data
Arrearage <- read.csv(here::here("Data_raw/public-arrearage-program-data-2022-01-19.csv")) #data comes from SWRCB (https://www.waterboards.ca.gov/arrearage_payment_program/). This 1-19-22 version has remained the most updated on their website, I think applications were due December 2021 so that makes sense?
Arrearage <- rename(Arrearage, PWSID = Water.System.ID....PWSID.)
Arrearage$PWSID <- as.factor(Arrearage$PWSID)
Arrearage <- Arrearage %>% select(PWSID, Intend.to.apply., Application.complete.)
Arrearage$Intend.to.apply. <- as.factor(Arrearage$Intend.to.apply.)
Arrearage$Application.complete. <- as.factor(Arrearage$Application.complete.)

SAFER2023 <- read.csv(here::here("Data_raw/Drinking_Water_Risk_Assessment.csv"))#SAFER data comes from https://data.ca.gov/dataset/safer-failing-and-at-risk-drinking-water-systems. Dictionary saved in data folder
SAFER2023 <- rename(SAFER2023, PWSID = WATER_SYSTEM_NUMBER)
SAFER2023 <- SAFER2023 %>% select(PWSID, SERVICE_CONNECTIONS, POPULATION, MHI, CALENVIRO_SCREEN_SCORE, FINAL_SAFER_STATUS, PRIMARY_MCL_VIOLATION, SECONDARY_MCL_VIOLATION,E_COLI_VIOLATION, TREATMENT_TECHNIQUE_VIOLATION, MONITORING_AND_REPORTING_VIOLATION, WEIGHTED_WATER_QUALITY_SCORE, WEIGHTED_ACCESSIBILITY_SCORE, WEIGHTED_AFFORDABILITY_SCORE, WEIGHTED_TMF_CAPACITY_SCORE, FUNDING_RECEIVED_SINCE_2017, TOTAL_WEIGHTED_RISK_SCORE_BEFORE_DIVIDING_BY_CATEGORY_COUNT, CURRENT_FAILING)
SAFER2023$WEIGHTED_ACCESSIBILITY_SCORE <- as.numeric(SAFER2023$WEIGHTED_ACCESSIBILITY_SCORE)
SAFER2023$WEIGHTED_TMF_CAPACITY_SCORE <- as.numeric(SAFER2023$WEIGHTED_TMF_CAPACITY_SCORE)
SAFER2023$WEIGHTED_AFFORDABILITY_SCORE <- as.numeric(SAFER2023$WEIGHTED_AFFORDABILITY_SCORE)
SAFER2023$WEIGHTED_WATER_QUALITY_SCORE <- as.numeric(SAFER2023$WEIGHTED_WATER_QUALITY_SCORE)
SAFER2023$FINAL_SAFER_STATUS <- as.factor(SAFER2023$FINAL_SAFER_STATUS)
SAFER2023$PWSID <- as.factor(SAFER2023$PWSID)
SAFER2023$MHI <- as.numeric(SAFER2023$MHI)
SAFER2023$CALENVIRO_SCREEN_SCORE <- as.numeric(SAFER2023$CALENVIRO_SCREEN_SCORE)
SAFER2023$TOTAL_WEIGHTED_RISK_SCORE_BEFORE_DIVIDING_BY_CATEGORY_COUNT[SAFER2023$TOTAL_WEIGHTED_RISK_SCORE_BEFORE_DIVIDING_BY_CATEGORY_COUNT == "Not Assessed"] <- NA
SAFER2023$TOTAL_WEIGHTED_RISK_SCORE_BEFORE_DIVIDING_BY_CATEGORY_COUNT <- as.numeric(SAFER2023$TOTAL_WEIGHTED_RISK_SCORE_BEFORE_DIVIDING_BY_CATEGORY_COUNT)
SAFER2023$CURRENT_FAILING <- as.factor(SAFER2023$CURRENT_FAILING)

## join all together and final type corrections
Data <- left_join(CWSgov, SAFER2023)
Data <- left_join(Data, Arrearage)

Data$FundingYorN <- ifelse(Data$FUNDING_RECEIVED_SINCE_2017 > 0, "Yes", "No")
Data$FundingYorN <- as.factor(Data$FundingYorN)
Data <- rename(Data, Total_risk = TOTAL_WEIGHTED_RISK_SCORE_BEFORE_DIVIDING_BY_CATEGORY_COUNT)

```

California has `r nrow(CWSgov)` Community Water Systems (CWS) (excluding CWS on Tribal lands regulated directly by EPA region 9). These systems operate under a wide diversity of governance arrangements. Governance type, in turn, corresponds with a variety of important institutional considerations including how decisions are made, the structure (and even existence) of the governing board, and how the governing board is selected and by who. This analysis explores how particularly the later consideration may be associated with performance of these systems as it pertains to California's Human Right to Water Law (AB 685). 

### Construction of independent variable of interest 

The proposed independent variable  of interest to test this question is an ordered factor with four possible options: 1) no governing board; 2) governing board not selected by residents served; 3) board selected by a limited group of residents served (e.g. landowners); and 4) board selected by all registered voters served. The category assigned to each of the 28 institutional types identified in the 2023 institutional update is displayed below. An assignment of "variable" indicates where board selection is variable per California Code. Each of these systems (roughly 90) will be hand coded for the final analysis. Finally, Tribal government institutional types and the unknown private system types are labeled as NA and excluded from consideration in all analyses. The later due to lack of information and the former as outliers due to their unique political structures as sovereign nations.

```{r tabular overview of draft IVs, include=TRUE, echo=FALSE}
Table1 <- CWSgov %>% select(Final_inst_update, boardselection) %>% distinct()
knitr::kable(Table1, col.names = c("2023 Institution types", "Board selection"))

```

#### Questions for feedback: What do you think about this variable and its construction? Is it an accurate and meaningful portrayal of board selection? What changes would you suggest? Would it be better to look at eligibility to serve on the board? 

### Proposed outcome variables

System performance can be measured in a variety of ways and there is reason to believe that certain elements of performance may be more influenced by accountability and representation that others. For this reason I propose to use three distinct measures of performance: 1) whether a water system is on California's failing systems list (binary variable, Y/N); 2) System Needs Assessment risk score (numeric variable, can be disagregated to look at the needs assessment components individually e.g. water quality or affordability); 3) whether the water system applied for COVID-19 arrearage relief on behalf of their customers. These outcome measures are summarized below. 


```{r tabular overview of DVs, include=TRUE, echo=FALSE, message=FALSE, warning=FALSE}
Table2 <- as.data.frame(summary(Data$CURRENT_FAILING))
Table2$Percent <- (Table2$`summary(Data$CURRENT_FAILING)`)/(sum(Table2$`summary(Data$CURRENT_FAILING)`))
Table2$Percent <- round(x = Table2$Percent, digits = 2)
knitr::kable(Table2, col.names = c("Count", "Percent of all systems"))

library(vtable)
Riskscore_and_components <- Data %>% select(Total_risk, WEIGHTED_WATER_QUALITY_SCORE, WEIGHTED_ACCESSIBILITY_SCORE, WEIGHTED_AFFORDABILITY_SCORE, WEIGHTED_TMF_CAPACITY_SCORE)
st(Riskscore_and_components, out = "kable", title = "Summary statistics for risk score and components")

Arrearage <- Data %>% select(Application.complete.)
st(Arrearage, out = "kable", title = "Summary of arrearage applications")
```


#### Questions for feedback: Are there other outcome or performance variables that would be interesting to  look at (e.g. total funding received in recent years)? How can we ensure a "well rounded" assessment of water system performance as it relates to the human right to water and best hone in on the influence of governance?  

### Proposed controls/covariates

To keep things simple, I'm thinking of using thee controls/covariates in the analyses: 1) population served (log transformed); 2) water source (groundwater or surface water); and 3) whether water is purchased or self-produced. 

#### Questions for feedback: Are there other controls that should be considered? How do we think about the role of geography? Is it necessary to include the purchased water variable?   

```{r add in controls, include=FALSE, echo=FALSE}
#population from safer
#water source and whether purchased from CA SDWIS download July 5
SDWIS <- read.csv(here::here("Data_raw/CA_SDWIS_alldownload_July52023.csv"))
SDWIS <- rename(SDWIS, PWSID = Water.System.No.)
SDWIS$PWSID <- as.factor(SDWIS$PWSID)
SDWIS$Type <- as.factor(SDWIS$Type)
SDWIS$Primary.Source.Water.Type <- as.factor(SDWIS$Primary.Source.Water.Type)
SDWIS <- SDWIS %>% select(PWSID, Type, Primary.Source.Water.Type)
SDWIS$Source <- as.factor(ifelse(SDWIS$Primary.Source.Water.Type == "GU" | SDWIS$Primary.Source.Water.Type == "GUP" | SDWIS$Primary.Source.Water.Type == "GW" | SDWIS$Primary.Source.Water.Type == "GWP", "GW", "SW"))
SDWIS$Purchased <- as.factor(ifelse(SDWIS$Primary.Source.Water.Type == "GUP" | SDWIS$Primary.Source.Water.Type == "SWP" | SDWIS$Primary.Source.Water.Type == "GWP", "Purchased", "Self-produced"))

Data <- left_join(Data, SDWIS)
Data$LN_POP <- log((Data$POPULATION+1)) #added one to population first to avoide issue of loging 0. Need to figure out if this is the best way....
```

### Exploratory results

I include the following "rough" regressions to help illustrate what this analysis could look like, results will change as the variables and data are finalized. 

```{r analyses, include=TRUE, echo=FALSE}
library(sjPlot)

Data$CURRENT_FAILING <- relevel(Data$CURRENT_FAILING, ref = "Not Failing")
Data$boardselection <- relevel(Data$boardselection, ref = "No board")
Failing <- glm(CURRENT_FAILING ~ boardselection + LN_POP + Source, data = Data, family= binomial)
summary(Failing)

Plot1 <- plot_model(Failing, type = "pred", terms = "boardselection", axis.title = c("Board selection", "Failing system status"), title = " ") + theme_sjplot(base_size = 10); Plot1

Risk <- lm(Total_risk ~ boardselection + LN_POP + Source, Data)
summary(Risk)

Plot2 <- plot_model(Risk, type = "pred", terms = "boardselection", axis.title = c("Board selection", "Predicted total risk score"), title = " ") + theme_sjplot(base_size = 10); Plot2

Data$Application.complete. <- relevel(Data$Application.complete., ref = "No")
COVID <- glm(Application.complete. ~ voting3 + LN_POP + Source, data = Data, family= binomial)
summary(COVID)

Plot3 <- plot_model(COVID, type = "pred", terms = "voting3", axis.title = c("Board selection", "Applied for covid arrearage relief"), title = " ") + theme_sjplot(base_size = 10); Plot3

```
