---
title: 'Voting rights and the HR2W: Initial analysis and next steps brainstorm'
author: "Kristin Dobbin"
date: "2023-06-23"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE, echo=FALSE}
#markdown settings
knitr::opts_knit$set(root.dir = "/Users/KristinDobbin/Dropbox/Projects/R_Projects/Votingrights_-_HR2W", warning=FALSE, message=FALSE, include=FALSE)

#load libraries
library(tidyverse)
```

```{r data, include=FALSE, echo=FALSE}
###Load data and initial manipulations

CWSgov <- read.csv(("Data/FINAL 2023 Inst. type update.csv"))
CWSgov$Primacy_FINAL <- as.factor(CWSgov$Primacy_FINAL)
CWSgov <- CWSgov %>% filter(Primacy_FINAL != "EPA Region 9") #for purposes of this project get rid of EPA regulated Tribal water systems
CWSgov$Primacy_FINAL <- droplevels(CWSgov$Primacy_FINAL)
CWSgov <- CWSgov[,c(1:4,10:11)]
CWSgov$PWSID <- as.factor(CWSgov$PWSID)
CWSgov$Counties_Served <- as.factor(CWSgov$Counties_Served)
CWSgov$Final_inst_update <- as.factor(CWSgov$Final_inst_update)
CWSgov$Inst_Subtype  <- as.factor(CWSgov$Inst_Subtype)

### create new variables
## create ancillary variable
CWSgov$ancillary <- ifelse(CWSgov$Final_inst_update == "Community Services District" | CWSgov$Final_inst_update == "County Service Area" | CWSgov$Final_inst_update == "Municipal Water District" | CWSgov$Final_inst_update == "City" | CWSgov$Final_inst_update == "County Sanitation District"| CWSgov$Final_inst_update == "County Water District"| CWSgov$Final_inst_update == "County Waterworks District" | CWSgov$Final_inst_update == "Irrigation District"| CWSgov$Final_inst_update == "Maintenance District"| CWSgov$Final_inst_update == "Municipal Utility District"| CWSgov$Final_inst_update == "Public Utility District"| CWSgov$Final_inst_update == "Resort Improvement District"| CWSgov$Final_inst_update == "Resource Conservation District"| CWSgov$Final_inst_update == "Sanitary District"| CWSgov$Final_inst_update == "Water Conservation District"| CWSgov$Final_inst_update == "Special Act District" | CWSgov$Final_inst_update == "Mutual Benefit" | CWSgov$Final_inst_update == "Joint Powers Authority/Agreement" | CWSgov$Final_inst_update == "Investor Owned Utility" | CWSgov$Final_inst_update == "California Water District","No", NA)

CWSgov$ancillary <- ifelse(CWSgov$Final_inst_update == "County" | CWSgov$Final_inst_update == "Federal" | CWSgov$Final_inst_update == "Private - Ancillary" | CWSgov$Final_inst_update == "School District" | CWSgov$Final_inst_update == "State" | CWSgov$Final_inst_update == "Mobile Home Park" , "Yes", CWSgov$ancillary)

## create public/private variable
CWSgov$public <- ifelse(CWSgov$Final_inst_update == "Community Services District" | CWSgov$Final_inst_update == "County Service Area" | CWSgov$Final_inst_update == "Municipal Water District" | CWSgov$Final_inst_update == "City" | CWSgov$Final_inst_update == "County Sanitation District"| CWSgov$Final_inst_update == "County Water District"| CWSgov$Final_inst_update == "County Waterworks District" | CWSgov$Final_inst_update == "Irrigation District"| CWSgov$Final_inst_update == "Maintenance District"| CWSgov$Final_inst_update == "Municipal Utility District"| CWSgov$Final_inst_update == "Public Utility District"| CWSgov$Final_inst_update == "Resort Improvement District"| CWSgov$Final_inst_update == "Resource Conservation District"| CWSgov$Final_inst_update == "Sanitary District"| CWSgov$Final_inst_update == "Water Conservation District"| CWSgov$Final_inst_update == "Special Act District" | CWSgov$Final_inst_update == "Mutual Benefit" | CWSgov$Final_inst_update == "Joint Powers Authority/Agreement" | CWSgov$Final_inst_update == "California Water District" | CWSgov$Final_inst_update == "County" | CWSgov$Final_inst_update == "Federal" | CWSgov$Final_inst_update == "School District" | CWSgov$Final_inst_update == "State" | CWSgov$Final_inst_update == "Tribal", "Yes", "No")

## create voting enfranchisement variable
CWSgov$voting <- ifelse(CWSgov$Final_inst_update == "Community Services District" | CWSgov$Final_inst_update == "County Service Area" | CWSgov$Final_inst_update == "Municipal Water District" | CWSgov$Final_inst_update == "City" | CWSgov$Final_inst_update == "County Sanitation District"| CWSgov$Final_inst_update == "County Water District"| CWSgov$Final_inst_update == "County Waterworks District" | CWSgov$Final_inst_update == "Irrigation District"| CWSgov$Final_inst_update == "Maintenance District"| CWSgov$Final_inst_update == "Municipal Utility District"| CWSgov$Final_inst_update == "Public Utility District"| CWSgov$Final_inst_update == "Resort Improvement District"| CWSgov$Final_inst_update == "Resource Conservation District"| CWSgov$Final_inst_update == "Sanitary District"| CWSgov$Final_inst_update == "Water Conservation District", "Full", NA)

CWSgov$voting <- ifelse(CWSgov$Final_inst_update == "Mutual Benefit" | CWSgov$Final_inst_update == "Irrigation District","Limited", CWSgov$voting)

CWSgov$voting <- ifelse(CWSgov$Final_inst_update == "Investor Owned Utility", "None", CWSgov$voting)

CWSgov$voting <- ifelse(CWSgov$Final_inst_update == "California Water District" | CWSgov$Final_inst_update == "Joint Powers Authority/Agreement" | CWSgov$Final_inst_update == "Special Act District" , "Variable", CWSgov$voting)

#create second version of voting variable including MHPs as "none"
CWSgov$voting2 <- ifelse(CWSgov$Final_inst_update == "Community Services District" | CWSgov$Final_inst_update == "County Service Area" | CWSgov$Final_inst_update == "Municipal Water District" | CWSgov$Final_inst_update == "City" | CWSgov$Final_inst_update == "County Sanitation District"| CWSgov$Final_inst_update == "County Water District"| CWSgov$Final_inst_update == "County Waterworks District" | CWSgov$Final_inst_update == "Irrigation District"| CWSgov$Final_inst_update == "Maintenance District"| CWSgov$Final_inst_update == "Municipal Utility District"| CWSgov$Final_inst_update == "Public Utility District"| CWSgov$Final_inst_update == "Resort Improvement District"| CWSgov$Final_inst_update == "Resource Conservation District"| CWSgov$Final_inst_update == "Sanitary District"| CWSgov$Final_inst_update == "Water Conservation District", "Full", NA)

CWSgov$voting2 <- ifelse(CWSgov$Final_inst_update == "Mutual Benefit" | CWSgov$Final_inst_update == "Irrigation District","Limited", CWSgov$voting2)

CWSgov$voting2 <- ifelse(CWSgov$Final_inst_update == "Investor Owned Utility" | CWSgov$Final_inst_update == "Mobile Home Park", "None", CWSgov$voting2)

CWSgov$voting2 <- ifelse(CWSgov$Final_inst_update == "California Water District" | CWSgov$Final_inst_update == "Joint Powers Authority/Agreement" | CWSgov$Final_inst_update == "Special Act District" , "Variable", CWSgov$voting2)

#create third version of voting variable including all ancillary systems as "none"
CWSgov$voting3 <- ifelse(CWSgov$Final_inst_update == "Community Services District" | CWSgov$Final_inst_update == "County Service Area" | CWSgov$Final_inst_update == "Municipal Water District" | CWSgov$Final_inst_update == "City" | CWSgov$Final_inst_update == "County Sanitation District"| CWSgov$Final_inst_update == "County Water District"| CWSgov$Final_inst_update == "County Waterworks District" | CWSgov$Final_inst_update == "Irrigation District"| CWSgov$Final_inst_update == "Maintenance District"| CWSgov$Final_inst_update == "Municipal Utility District"| CWSgov$Final_inst_update == "Public Utility District"| CWSgov$Final_inst_update == "Resort Improvement District"| CWSgov$Final_inst_update == "Resource Conservation District"| CWSgov$Final_inst_update == "Sanitary District"| CWSgov$Final_inst_update == "Water Conservation District", "Full", NA)

CWSgov$voting3 <- ifelse(CWSgov$Final_inst_update == "Mutual Benefit" | CWSgov$Final_inst_update == "Irrigation District","Limited", CWSgov$voting3)

CWSgov$voting3 <- ifelse(CWSgov$Final_inst_update == "Investor Owned Utility" | CWSgov$Final_inst_update == "Mobile Home Park" | CWSgov$Final_inst_update == "County" | CWSgov$Final_inst_update == "Federal" | CWSgov$Final_inst_update == "Private - Ancillary" | CWSgov$Final_inst_update == "School District" | CWSgov$Final_inst_update == "State", "None", CWSgov$voting3)

CWSgov$voting3 <- ifelse(CWSgov$Final_inst_update == "California Water District" | CWSgov$Final_inst_update == "Joint Powers Authority/Agreement" | CWSgov$Final_inst_update == "Special Act District" , "Variable", CWSgov$voting3)
##mutate to create new variables following initial script but instead of NA for variable, use variable and can filter out later but that way can check with Camille and Justin and discuss what to do

## create final variable about whether there is a board and if so who is elgible to serve on it
#create third version of voting variable including all ancillary systems as "none"
CWSgov$board <- ifelse(CWSgov$Final_inst_update == "Community Services District" | CWSgov$Final_inst_update == "County Service Area" | CWSgov$Final_inst_update == "Municipal Water District" | CWSgov$Final_inst_update == "City" | CWSgov$Final_inst_update == "County Sanitation District"| CWSgov$Final_inst_update == "County Water District"| CWSgov$Final_inst_update == "County Waterworks District" | CWSgov$Final_inst_update == "Irrigation District"| CWSgov$Final_inst_update == "Maintenance District"| CWSgov$Final_inst_update == "Municipal Utility District"| CWSgov$Final_inst_update == "Public Utility District"| CWSgov$Final_inst_update == "Resort Improvement District"| CWSgov$Final_inst_update == "Resource Conservation District"| CWSgov$Final_inst_update == "Sanitary District"| CWSgov$Final_inst_update == "Water Conservation District", "Full", NA)

CWSgov$board <- ifelse(CWSgov$Final_inst_update == "California Water District" | CWSgov$Final_inst_update == "Mutual Benefit" | CWSgov$Final_inst_update == "Irrigation District","Limited", CWSgov$board)

CWSgov$board <- ifelse(CWSgov$Final_inst_update == "Investor Owned Utility" | CWSgov$Final_inst_update == "Mobile Home Park" | CWSgov$Final_inst_update == "County" | CWSgov$Final_inst_update == "Federal" | CWSgov$Final_inst_update == "Private - Ancillary" | CWSgov$Final_inst_update == "School District" | CWSgov$Final_inst_update == "State", "None", CWSgov$board)

CWSgov$board <- ifelse( CWSgov$Final_inst_update == "Joint Powers Authority/Agreement" | CWSgov$Final_inst_update == "Special Act District" | CWSgov$Final_inst_update == "County Sanitation District" | CWSgov$Final_inst_update == "Irrigation District", "Variable", CWSgov$board)

#change to factors
CWSgov$ancillary <- as.factor(CWSgov$ancillary)
CWSgov$public <- as.factor(CWSgov$public)
CWSgov$voting <- as.factor(CWSgov$voting)
CWSgov$voting2 <- as.factor(CWSgov$voting2)
CWSgov$voting3 <- as.factor(CWSgov$voting3)

## add in outcome data
Arrearage <- read.csv(("Data/public-arrearage-program-data-2022-01-19.csv")) #data comes from SWRCB (https://www.waterboards.ca.gov/arrearage_payment_program/). This 1-19-22 version has remained the most updated on their website, I think applications were due December 2021 so that makes sense?
Arrearage <- rename(Arrearage, PWSID = Water.System.ID....PWSID.)
Arrearage$PWSID <- as.factor(Arrearage$PWSID)
Arrearage <- Arrearage %>% select(PWSID, Intend.to.apply., Application.complete.)
Arrearage$Intend.to.apply. <- as.factor(Arrearage$Intend.to.apply.)
Arrearage$Application.complete. <- as.factor(Arrearage$Application.complete.)

SAFER2023 <- read.csv("Data/Drinking_Water_Risk_Assessment.csv") #SAFER data comes from https://data.ca.gov/dataset/safer-failing-and-at-risk-drinking-water-systems. Dictionary saved in data folder
SAFER2023 <- rename(SAFER2023, PWSID = WATER_SYSTEM_NUMBER)
SAFER2023 <- SAFER2023 %>% select(PWSID, SERVICE_CONNECTIONS, POPULATION, MHI, CALENVIRO_SCREEN_SCORE, FINAL_SAFER_STATUS, PRIMARY_MCL_VIOLATION, SECONDARY_MCL_VIOLATION,E_COLI_VIOLATION, TREATMENT_TECHNIQUE_VIOLATION, MONITORING_AND_REPORTING_VIOLATION, WEIGHTED_WATER_QUALITY_SCORE, WEIGHTED_ACCESSIBILITY_SCORE, WEIGHTED_AFFORDABILITY_SCORE, WEIGHTED_TMF_CAPACITY_SCORE, FUNDING_RECEIVED_SINCE_2017, TOTAL_WEIGHTED_RISK_SCORE_BEFORE_DIVIDING_BY_CATEGORY_COUNT)
SAFER2023$WEIGHTED_ACCESSIBILITY_SCORE <- as.numeric(SAFER2023$WEIGHTED_ACCESSIBILITY_SCORE)
SAFER2023$WEIGHTED_TMF_CAPACITY_SCORE <- as.numeric(SAFER2023$WEIGHTED_TMF_CAPACITY_SCORE)
SAFER2023$WEIGHTED_AFFORDABILITY_SCORE <- as.numeric(SAFER2023$WEIGHTED_AFFORDABILITY_SCORE)
SAFER2023$WEIGHTED_WATER_QUALITY_SCORE <- as.numeric(SAFER2023$WEIGHTED_WATER_QUALITY_SCORE)
SAFER2023$FINAL_SAFER_STATUS <- as.factor(SAFER2023$FINAL_SAFER_STATUS)
SAFER2023$PWSID <- as.factor(SAFER2023$PWSID)
SAFER2023$MHI <- as.numeric(SAFER2023$MHI)
SAFER2023$CALENVIRO_SCREEN_SCORE <- as.numeric(SAFER2023$CALENVIRO_SCREEN_SCORE)
SAFER2023$TOTAL_WEIGHTED_RISK_SCORE_BEFORE_DIVIDING_BY_CATEGORY_COUNT[SAFER2023$TOTAL_WEIGHTED_RISK_SCORE_BEFORE_DIVIDING_BY_CATEGORY_COUNT == "Not Assessed"] <- NA
SAFER2023$TOTAL_WEIGHTED_RISK_SCORE_BEFORE_DIVIDING_BY_CATEGORY_COUNT <- as.numeric(SAFER2023$TOTAL_WEIGHTED_RISK_SCORE_BEFORE_DIVIDING_BY_CATEGORY_COUNT)

## join all together and final type corrections
Data <- left_join(CWSgov, SAFER2023)
Data <- left_join(Data, Arrearage)

Data$FundingYorN <- ifelse(Data$FUNDING_RECEIVED_SINCE_2017 > 0, "Yes", "No")
Data$FundingYorN <- as.factor(Data$FundingYorN)
Data <- rename(Data, Total_risk = TOTAL_WEIGHTED_RISK_SCORE_BEFORE_DIVIDING_BY_CATEGORY_COUNT)
```

California has `r nrow(CWSgov)` Community Water Systems (CWS) (excluding CWS on Tribal lands regulated directly by EPA region 9). These systems operate under a wide diversity of governance arrangements. Governance type, in turn, corresponds with a variety of important institutional considerations including how decisions are made, the structure (and even existence) of the governing board, and how the governing board is selected and by who. This analysis explores how particularly the later consideration may be associated with performance of these systems as it pertains to California's Human Right to Water Law (AB 685). 

### Construction of independent variables of interest 

The proposed independent variable  of itnerest to test this question is an ordered factor with four possible options: 1) no governing board; 2) governing board not selected by residents served; 3) board selected by a limited group of residents served (e.g. landowners); and 4) board selected by all registered voters served. The category assigned to each of the 28 institutional types identified in the 2023 institutional update is displayed below.   

at these two independent variables of interest for all 28 institutional types identified in the 2023 update, with two versions of the voting enfranchisement variable displayed for consideration. The first voting variable ("voting"), excludes from consideration ("NA") institutional types without governing boards (generally ancillary systems that exist for a primary purpose other than water provision). For example, state and federal systems that are recreation and carceral facilities and private ancillary systems that are private businesses including farms, packing houses, industrial parks etc.  The remainder of types are classified as either having no resident voting for the governing board, limited resident voting (ie limited to landowners) or full resident voting (ie all registered voters in district can vote). Alternatively, the second version of the voting variable ("Voting3") includes ancillary systems without governing boards in the category of "none" for the purposes of characterizing resident enfranchisement (in other words, groups together systems without a board with those where there is a board but it is not elected). Either one of these two variables could be used in the analysis or a different approach altogether could be used (e.g. two variables, one for existence of the board and one for voter enfranchisement where there is a board).  

Finally, the "board" variable indicates the extent to which registered voter residents of the district are eligible to serve on the board themselves. Full means any registered voter can serve. Limited means only landowners can serve and none means there is either no board or the board is not comprised of local residents (so more similar to the voting3 variable). The big issue with this variable is that the public/governmental districts are limited to registered voters which necessitates citizenship which is not true of private types, most notable Mutual Benefit systems. It is hard to capture both of these considerations in one variable. 

In each of the three variables, Tribal government institutional types and the unknown private system types are excluded from consideration. The later due to lack of information and the former as outliers due to their unique political structures as sovereign nations. Institutional types where voting enfranchisement or board eligibility is variable per California Code are currently listed as "variable". My plan is to code these 90-100 systems individually. 

```{r tabular overview of draft IVs, include=TRUE, echo=FALSE}
Table1 <- CWSgov %>% select(Final_inst_update, voting, voting3, board) %>% distinct()
knitr::kable(Table1, col.names = c("2023 Institution types", "Voting", "Voting3", "Board"))

```

#### Questions for feedback: What do you think about these three variables? How could they be improved? Which is most compelling/informative in terms of an anlysis of system performance? Do you disagree wtih any of the characterizations or see any mistakes?

### Potential system performance/outcome variables

System performance, which can be measured in a variety of ways, is similarly diverse. California's CWSs vary with respect to their SAFER Needs Assessment status which reflects different levels of risk across water quality, water accessibility, water affordability and Technical Managerial and Financial Capacity. 

```{r tabular overview of DVs, include=TRUE, echo=FALSE}
Table2 <- as.data.frame(summary(Data$FINAL_SAFER_STATUS))
Table2$Percent <- (Table2$`summary(Data$FINAL_SAFER_STATUS)`)/(sum(Table2$`summary(Data$FINAL_SAFER_STATUS)`))
Table2$Percent <- round(x = Table2$Percent, digits = 2)
knitr::kable(Table2, col.names = c("Count", "Percent of all systems"))

knitr::kable(summarize(Data,
           Mean_total_risk = mean(na.omit(Total_risk)),
           Mean_quality_risk = mean(na.omit(WEIGHTED_WATER_QUALITY_SCORE)),
           Mean_accessibility_risk = mean(na.omit(WEIGHTED_ACCESSIBILITY_SCORE)),
           Mean_affordability_risk = mean(na.omit(WEIGHTED_AFFORDABILITY_SCORE)),
           Mean_TMF_risk = mean(na.omit(WEIGHTED_TMF_CAPACITY_SCORE))))

```

Notably, system risk or failing status reflects a complicated combination of factors and causes, not all of which are necessarily within the control of the governing board. Even assuming that voting and board representation matters, it is entirely feasible that a water system with robust representation and community participation is governance might still struggle if it is very small, has poor source water quality, has dilapidated infrastructure etc. Some additional potential outcomes for variation that may more clearly reflect the role of governance in system performance include funding received (likely still reflects risk some but also potentially reflects proactive efforts to fix issues?). `r sum(na.omit(Data$FUNDING_RECEIVED_SINCE_2017>0))` of CWSs have received state funding since 2017. Another potential option is whether a system applied for COVID-19 arrearage relief. We have data both on whether a system reported to the SWRCB that they intended to apply for the funding and if they did.  

```{r more tabular overview of DVs, include=TRUE, echo=FALSE}
Table3 <- as.data.frame(summary(Data$Intend.to.apply.))
Table3$Percent <- (Table3$`summary(Data$Intend.to.apply.)`)/(sum(Table3$`summary(Data$Intend.to.apply.)`))
Table3$Percent <- round(x = Table3$Percent, digits = 2)
knitr::kable(Table3, col.names = c("Reported intending to apply", "Percent of all systems"))

Table4 <- as.data.frame(summary(Data$Application.complete.))
Table4$Percent <- (Table4$`summary(Data$Application.complete.)`)/(sum(Table4$`summary(Data$Application.complete.)`))
Table4$Percent <- round(x = Table4$Percent, digits = 2)
knitr::kable(Table4, col.names = c("Applied", "Percent of all systems"))
```

#### Questions for feedback: Are there other outcome or performance variables that would be interesting to  look at? How can we ensure a "well rounded" assessment of water system performance as it relates to the human right to water and best hone in on the influence of governance?  


### Exploratory results

Once the independent variables and outcomes of interest have been solidified, I will run some formal analyses using more appropriate statistical techniques but I include the following "rough" linear regressions to help illustrate what this analysis could look like (note covariates/controls currently not included, I plan to use at least service connections (or population) and source water, could include county).

```{r analyses, include=TRUE, echo=FALSE}
Vote1 <- lm(Total_risk ~ voting3, Data)
summary(Vote1)

library(sjPlot)
Plot1 <- plot_model(Vote1, type = "pred", terms = "voting3", axis.title = c("Resident enfranchisement", "Predicted risk score"), title = " ") + theme_sjplot(base_size = 18); Plot1

Vote2 <- glm(Application.complete. ~ voting3, data = Data, family= binomial)
summary(Vote2)

Plot2 <- plot_model(Vote2, type = "pred", terms = "voting3", axis.title = c("Resident enfranchisement", "Applied for arrearage $"), title = " ") + theme_sjplot(base_size = 18); Plot2

Board1 <- lm(Total_risk ~ board, Data)
summary(Board1)

Plot3 <- plot_model(Board1, type = "pred", terms = "board", axis.title = c("Board eligibility", "Predicted risk score"), title = " ") + theme_sjplot(base_size = 18); Plot3

Board2 <- glm(FundingYorN ~ board, data = Data, family= binomial)
summary(Board2)

Plot4 <- plot_model(Board2, type = "pred", terms = "board", axis.title = c("Board eligibility", "Applied for state $ since 2017"), title = " ") + theme_sjplot(base_size = 18); Plot4
```


```{r tabular overview of covariates, include=FALSE, echo=FALSE, eval=FALSE}

knitr::kable(summarize(Data,
           Count = n(),
           Mean_population = mean(na.omit(POPULATION)),
           Mean_serviceconnections= mean(na.omit(SERVICE_CONNECTIONS)),
           Mean_MHI = mean(na.omit(MHI)),
           Mean_Calenviroscreen = mean(na.omit(CALENVIRO_SCREEN_SCORE))))
```


